\subsection{Robotics}
\label{sec:robotics}
In may 2023 I asked chatgbt to give me some applications of the Gauss-Bonnet theorem and returned
that there were applications to robotics. I asked for some references here is what is said
``Yes, here are some references for the applications of the discrete Gauss-Bonnet theorem to robotics: 1. “Geometric Mechanics Approach to Control of Superparamagnetic Beads in High-Throughput Robotic Platforms” by P. Szymczak, E. Garcia-Miralles, and S. Misra, IEEE/ASME Transactions on Mechatronics (2015). 2. “Geometric Path Planning for Robotics Using Discrete Gauss-Bonnet Theorem” by V. Sarma and B. Ravani, IEEE International Conference on Robotics and Automation (2011). 3. “Gaussian Curvature-Based Navigation Algorithm for Mobile Robots” by Z. Chen, X. Yan, and Y. Zhang, IEEE Transactions on Robotics (2012). 4. “A Curvature-Based Approach to Visual Servoing for Robotic Manipulators” by Y. Liu and G. Hu, IEEE Transactions on Robotics (2010). These references provide examples of how the discrete Gauss-Bonnet theorem and related geometric techniques can be used in robotics, including control of magnetic beads, geometric path planning, navigation algorithms, and visual servoing."

These references seem to be made up but some web searches lead me to the following
not made up application of the Gauss-Bonnet theorem in robotic 
route planning is given by K.-L. Wu et. al. in \cite{wu_path_2016}.
Suppose have a robot navigating a 3D\todo{2 or 3 manifold?} terrain with a single obstacle
and we wish to plan trips for our robot.
In this application, a terrain is a smooth manifold \todo{not defined} with tangent planes
at every point. An obstacle is modeled by a hazardous ball with a grade depending on the radius.
Assume that checking if a path intersects the obstacle can be done in constant time.


Overview of their procedure.
We are given two points $s$ and $t$ on a ?-manifold $M$.
First, compute a geodesic path from  $s$ to $t$ call this path $\gamma(x)$.
If the $\gamma$ does not intersect the obstacle we are done.
Otherwise, the $\gamma$ intersects the obstacle
we call the initial intersection point between our path
and the boundary of the obstacle $p$.
Construct the tanged plant $TpM$ at $p$.
Choose a vector $v\in TpM$ and a value $\alpha$ for the magnitude of $v$.
Next, define two points $\alpha_{\ell}$ and $\alpha_{u}$
to be in the directions of $v$ and $-v$ at a distance of $\alpha$
from $p$ in the tangent plane. Then project the points   $\alpha_{\ell}$ and $\alpha_{u}$
onto the surface? to obtain the points $q$ and $r$.
We next compute four new  geodesics $g_1(s)$ from $s$ to $q$,
$g_2(s)$ from $q$ to $t$, 
$f_1(s)$ from $s$ to $r$ and 
$f_2(s)$ from $r$ to $t$. Let $\gamma_g=g_2\circ g_1$ and $\gamma_f=f_2\circ f_1$.

We then use the Gauss-Bonnet theorem to decide which alternative path is best.
If the edges of a triangle are all geodesic then we have a \EMPH{geodesic triangle} $\tau\subset M$.
We have two geodesic triangles, $\gamma, g_1,g_2$ and $\gamma,f_1,f_2.$

Here we have cusps at the intersection points of the geodesics,
to account for this, our Gauss-Bonnet is 
\begin{equation}\label{eqn:b-g-angles}
\int \int_{\tau} K dA +\sum_{i=1}^3(\pi-\theta_i)+\sum_{i=1}^3 \int k_gds =2\pi
\end{equation}
where $\theta_i$ are the interior angles of $\tau$.
Since we are on geodesics $\int kgds =0$ and we can rearrange
\eqnref{b-g-angles} to obtain

\begin{equation} \label{eqn:interior-angles}
\theta_1+\theta_2+\theta_3 = \pi +\int \int_{\tau} K dA.
\end{equation}

We can estimate the curvature based on the sum of the angles of
$\sum_{i=1}3\theta$.
They show that f $K=0$ on $\tau$ then $\gamma_f$ and $\gamma_g$
are identical and there exists an $\alpha^*$ that makes them shortest.
If $\sum_{i=1}3\theta_i>\pi$ then is $\int_?K>0$ on all of $\tau$ then and if $\sum_{i=1}3\theta_i<\pi$
then $K<0$ on (average) all of $\tau$.
The authors state that we should avoid negative curvature because it
is ``energy-consuming ascending and descending motions required"
and would require the robot have ``better mobility and maneuverability".
\todo{I don't see why}.


